PhD Title \\ PhD Thesis

### 2024-07-24 Latex_files/Thesis/chapters/chapter 2/main.tex
updates

Optical tweezing operates on the principle that light carries both
linear and angular momentum, which can be transferred to other
objects. If light is reflected or refracted by a medium part of the
light's momentum is transferred to the medium itself. While at large
scales this effect is insignificant in most cases, at micron scales
the forces imparted by individual photons begin to become significant
enough to influence its trajectory. Ashkin's work into optical
tweezing found that micron sized entities could be spatially locked by
aiming a laser with a Gaussian intensity profile directly upward at a
the cell while being suspended in a liquid medium \cite{Ashkin1970}.
This is demonstrated diagram below where we have a simple sphere
trapped by a Gaussian beam, if we consider two rays of light hitting
our sphere we see that while in the centre each ray is refracted in an
equal opposite direction to one another and so the net force imparted
is 0.  However when displaced the net force is unbalanced and points
towards the centre of the trap; this principle can be generalised to 3
degrees of freedom if we consider a focused beam instead of a
para-axial beam, typically the trap will be weaker along the beam axis
compared to the transverse trap strength.

\begin{figure}[h]
  \centering
  \includegraphics[width=0.65\linewidth]{ot_working_principal.png}
  \caption{Working principle for an optical trap, upper image shows a sphere at the centre of a trap experiences equal forces directed towards inwards. Lower image shows that when displaced 'ray a' is refracted far more than 'ray b' resulting in a net force back towards the centre.}
\end{figure}

@@

These 4 equations describe how the electric and magnetic fields behave
and are related to one another at a microscopic level. Any discussion 
of optical trapping is underpinned by the fact that in all cases the 
Maxwell Equations must be satisfied at the macroscopic level, where 
one must account for the medium's interactions with the EM field. 
Macroscopic forms of the maxwell equations are given below \cite{Jackson_1975}:
\begin{align}
	\nabla \cdot \mathbf{D}
	&= \rho_f
	\\
	\nabla \cdot \mathbf{B}
	&= 0
	\\
	\nabla \times \mathbf{E}
	&= \frac{\partial \mathbf{B}}{\partial t}
	\\
	\nabla \times \mathbf{H}
	&= \mathbf{J}_f +\frac{\partial \mathbf{D}}{\partial t}  
\end{align}
Where $\rho_f$ and $\mathbf{J}_f$ are the charge density and current 
caused by the presence of free charges in the medium, and $\mathbf{D}$ 
and $\mathbf{H}$ are the displacement and magnetizing fields respectively. 
The latter two are defined by:
\begin{align}
	\nonumber
	\mathbf{D} &= \epsilon_0\mathbf{E}+\mathbf{P} 
	\\ 
	\nonumber
	\mathbf{H} &= \frac{1}{\mu_0}\mathbf{B}-\mathbf{M}
\end{align}
Where $\mathbf{P}$ is the polarization field and $\mathbf{M}$ is the 
magnetisation field, these fields arise due to the bound charges 
throughout the medium interacting with the EM field.

The force exerted by an optical tweezer can be subdivided into the 
gradient and scattering components, for most modelling research this 
is how the force fields are reported. The gradient force is a conservative 
force brought about by the polarisation of dielectric materials, 
anything around 10\% of the trapping wavelength will be attracted along 
the gradient of the electric field to the region of greatest intensity 
(for a simple Gaussian Beam this would be at the focal point). The 
scattering force arises from the scattered field of the trapping beam 
pushing the target object away from the focal point; unlike the gradient \\
force, the scattering force is non-conservative and is loosely 
proportional to the particle's Brownian motion. The equilibrium position 
(where the mean squared displacement is minimised) is found when the 
gradient force far exceeds the scattering force, interpreting and 
calculating these forces is dependent on the ratio between the particle 
size and the trapping wavelength.

@@

The Ray-Optics model is the simplest to understand, the theory models 
the laser as a collection of individual 'rays' that propagate and are 
refracted according to Snell's Law. Based on the change in direction 
momentum is transferred to the target particle; with rays closest to 
the centre of the beam having greater intensity than those rays at 
the very edge of the beam. Consider a particle struck by two rays in 
a Gaussian beam (see Fig below), one coming close to the centre, and 
the other ray coming from the edge. As each ray is refracted by the 
target sphere a force is imparted onto it, the total force imparted 
is given by:
\begin{align}
	F_i = Q_i\frac{\Delta n P_i}{c}
\end{align}
where $Q_i$ is the trapping efficiency, $\Delta n$ is the difference
in refractive indices between the solution and the target particle,
and $P_i$ is the power of the individual ray. For a beam with a
Gaussian intensity distribution $P_i$ will fall off as you move from
the centre of the beam.  The ray optics model is ideal for

@@

Where $k$ is the wave number of the electromagnetic radiation ($k = 2\pi/\lambda$)
 This combined with the constraints of Maxwell's Equations leaves 
 very few exact solutions apart from spherical or planar waves, 
 both of which can be converted between with relative ease. For 
 example, a plane wave electric field can be expanded into spherical 
 harmonics and likewise any spherical wave can be described as 
 combination of plane waves offset from one another. For a single 
 plane wave the incident, internal, and scattered fields are given as:
\begin{align}
	\label{eq:lornez_mie}
  {\bf E}_{\rm inc}({\bf r})
  &=
    E_0 \sum_{nm}^\infty \left[
    a_{nm}{\bf M}_{nm}^{(1)}({\bf r})
    + b_{nm}{\bf N}_{nm}^{(1)}({\bf r})\right]
  \\
  \label{eq:internal}
  {\bf E}_{\rm int}({\bf r})
  &=
    E_0 \sum_{n=1}^\infty i^n\frac{2n+1}{n(n+1)}\left[
    -id_{nm}{\bf N}_{nm}^{(1)}({\bf r}) + c_{nm}{\bf M}_{nm}^{(1)}({\bf r})
    \right]
  \\
  {\bf E}_{\rm scat}({\bf r})
  &=
    E_0 \sum_{n=1}^\infty  i^n\frac{2n+1}{n(n+1)}\left[
      -i p_{nm}{\bf N}_{nm}^{(3)}({\bf r})+q_{nm}{\bf M}_{nm}^{(3)}({\bf r})
    \right] 
    \\
	\mathbf{E}_{tot}({\bf r}) &= 
	\begin{cases}
		{\bf E}_{\rm inc}({\bf r}) + {\bf E}_{\rm scat}({\bf r}) & \text{if \textbf{r} outside}
		\\ 
		{\bf E}_{\rm int}({\bf r}) & \text{if \textbf{r} inside}
	\end{cases}
\end{align}
where $a_{nm}$, $b_{nm}$, $c_{nm}$, $d_{nm}$, $p_{nm}$, and $q_{nm}$
are the expansion coefficients of each of the fields, and $M^{(1)}_{nm}$ 
and $N^{(1)}_{nm}$ are the vector spherical harmonics. For the incident
field computing its expansion coefficients is possible via analytical
methods and are completely dependent on the beam conditions imposed by
the user. However, computing the expansion coefficients of the
internal and external fields be far more tedious depending on the
shape of the target and what properties of the scattered field are
desired --- all of which is discussed later on.  From an optical
trapping perspective the force exerted by a focused electric field can
be found by computing the Maxwell stress tensor which only requires
the total magnitude of the incident and scattered fields, essentially
computing the momentum contained in the incident beam and how much of
it has been transferred to the target particle.

Lorenz-Mie theory can be applied to describe the scattering from any 
particle regardless of size, however as we either increase or decrease 
the size of the target particles the infinite series converges allowing 
one to ignore much of the tedium of Lorenz-Mie theory. The ray optics 
model is what is achieved when the size of the target particle far exceeds that of the laser wavelength and thus individual ray's can have independent contributions. 
In the latter case we can simplify the scattering problem by approximating 
our target as a single dipole, focusing only its interactions with the 
incident field.

@@

The Rayleigh approximation is the limiting approximation for describing 
a particles motion in an electromagnetic field who's wavelength is several 
times greater than the particle's size. The underlying theory is that a \\
dielectric sphere can be treated as a dipole while in the presence of the 
electromagnetic field; in which case the scattering force is given simply 
by the scattering of the induced dipole, and the gradient force is due 
to the Lorentz force \cite{Gordon1973}. The gradient forces in the 
principle Cartesian axis are described by Harada et al \cite{YasuhiroHarada1996} 
in MKS units as a restorative rectangular force field:
\begin{align}
  F_{grad,x}
  &=-\hat{x} \frac{2\pi n_2 a^3}{c}
    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{x}/w_0}{1+(2\tilde{z})^2} \times I(r)
  \\
  F_{grad,y}
  &=-\hat{y} \frac{2\pi n_2 a^3}{c}
    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{y}/w_0}{1+(2\tilde{z})^2} \times I(r)
  \\
  F_{grad,z}
  &=-\hat{z} \frac{2\pi n_2 a^3}{c}
    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{y}/w_0}{1+(2\tilde{z})^2}
    \nonumber \times I(r)
  \\ 
  & \times \left[1-\frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right]
  \\
  \text{where:}
  \nonumber
  \\
	I(r) &= \left(\frac{2P}{\pi w_0^2}\right) \frac{1}{1+(2\tilde{z}^2)} 
	\exp \left[ - \frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right]
\end{align}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Scattering methods}
\label{sec:scattering}
There are several methods available to calculate the scattered field 
produced by a particle, while one can compute this from directly applying
Lorenz-Mie theory for complicated particle shapes and focused beams makes it 
tedious for some applications.  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{T-matrix Method}
The $T$-matrix method was first developed by Peter Waterman with his
research into acoustic wave scattering \cite{Waterman1969}, this 
would later be extended to electromagnetic waves. Sometimes being 
referred to as the extended boundary condition method (ECBM), 
the method replaces the scatterer with a series of surface currents
over the targets surface, these currents are chosen so that the
electric field outside is identical to the original problem \cite{Wriedt1998}. 
This choice of currents negates the need to compute the internal fields
which reduces the scattering problem to a system of linear equations 
relating the incident beam coefficients to the scattered beam coefficients
the relationship between each can be summarised as:
\begin{align}
	\begin{pmatrix}
		q_{mn} \\
		p_{mn} 
	\end{pmatrix}
	= \bold{T} 
	\begin{pmatrix}
		a_{mn} \\
		b_{mn}
	\end{pmatrix}
	= \begin{pmatrix}
		T_{11} \ T_{12} \\
		T_{21} \ T_{22}
	\end{pmatrix}
	\begin{pmatrix}
		a_{mn} \\
		b_{mn}
	\end{pmatrix}
\end{align}

Where $q_{mn}$, and $p_{mn}$ are the scattering beam coefficients, and
$a_{mn}$, and $b_{mn}$ are the coefficients of the incident beam. For 
this PhD in particular we utilise an extension of the ECBM known 
as multi-sphere $T$-matrix method (\textit{mstm}), developed by Mackowski
\cite{Mackowski2011} the computational code computes the scattered field 
from each sphere within the target cluster and the incident field itself, 
this scattering between spheres converges to a final result but modern codes 
truncate the calculations to fit a desired accuracy and computational time.

The $T$-matrix method is exceptionally useful for computing the scattering 
from any arbitrary spherical aggregate. However, the $T$-matrix method by 
itself can be computationally taxing as the number of spheres increases 
\cite{Mackowski2011}. While it is possible to solve for the scattering of
each individual sphere this is only applicable for a single orientation and 
can be even slower for large aggregates \cite{Mackowski1996, Xu1995}. The 
benefit of \textit{mstm} is that the major scattering properties (scattering 
and extinction cross sections, scattering matrices, etc) can all be computed 
both for single orientations, or averaged over multiple orientations to determine
the average scattering from the target particle.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Discrete Dipole Approximation}
The discrete dipole approximation (DDA) is a general method that can 
be applied to the scattering from particles of arbitrary composition
and geometry. Developed by Purcell and Pennypacker \cite{Purcell1973}, 
the DDA method approximates the particle as being constructed of dipoles.
Each dipole interacts with both the incident field and the scattered fields
from every other dipole surrounding it. The resulting scattered field is 
identical to the scattered field produced by direct integration of 
Eq.~\ref{eq:internal} throughout the full particle volume \cite{Goedecke1988}.
The integral form for the internal electric field inside a scatterer is 
given as:
\begin{align}
	\mathbf{E(r)} = \mathbf{E_{inc}(r)} + \int_{V/V_0}d^3r\mathbf{\bar{G}(r,r')}
	\chi(\mathbf{r'})\mathbf{E(r')}
\end{align}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Langevin Equation}
Describing any microscopic motion requires an understanding of a molecules
diffusive behaviour, for the case of optical tweezers the most complete model 
of diffusion is the Langevin equation. Models such as the Fickian, and Einstein 
derivations are sufficient for macroscopic behaviours the Langevin equation 
better describes the microscopic characteristics of any diffusive behaviour. 
The Fickian model is not suitable for the applications of optical tweezers as 
it assumes that all diffusive motion is driven by a gradient of molecular density
\cite{Gillespie2012}, not only is this often not the case but it fails to describe 
the motion of any single molecule, only the collective behaviour. Einstein's model 
however considers the collisions experienced by individual molecules \cite{Gillespie2012a}, 
if we assume that all collisions are random and only consider the behaviour 
after a finite time step $\delta t$ then the individual collisions should 
cancel out, of course trying to understand the dynamics over a shorter period
of time is not described by Einstein's model as you would have to consider
each collision in turn \cite{Gillespie2012, Gillespie2012}. This is insufficient 
for our interests as an optical tweezers relaxation time (given by 
$\tau = \kappa/\gamma$) is often so short that we must consider the individual 
collisions experienced by any one molecule. The Langevin model of diffusion 
therefore assumes that the net force on a particular particle is described 
fully by these individual collisions \cite{Gillespie2012c}:
\begin{align}
	m\frac{dv}{dt} + \gamma_0 v + F(t) = W(t)
\end{align}

@@

Eq.~\ref{eq:langevin} provides an accurate description of strongly trapped 
particles, however the analytical solution of the Langevin equation requires 
integration of the white noise term making it difficult to simulate the 
trajectory of a given particle \cite{Volpe2013}. Instead it is often far 
easier to solve the equation numerically and apply use the analytical 
solution to calibrate and extract information about the particle and fluid, 
and how the two interact with the optical trap.

>>>>

Optical tweezing operates on the principle that light carries both
linear and angular momentum, which can be transferred to other
objects. If light is reflected or refracted by a medium part of the
light's momentum is transferred to the medium itself. While at large
scales this effect is insignificant in most cases, at micron scales
the forces imparted by individual photons begin to become significant
enough to influence the trajectory of individual particles. Ashkin's 
work into optical tweezing found that micron sized entities could be 
spatially locked by aiming a laser with a Gaussian intensity profile 
directly upward at a the cell while being suspended in a liquid medium 
\cite{Ashkin1970}.

This is demonstrated in fig.~\ref{fig:working_principle} where we have 
a simple sphere trapped by a circularly symmetrical beam with the intensity
falling off from the centre in a normal distribution (commonly referred to
as a Gaussian beam). If we consider two rays of light hitting our sphere 
we see that while in the centre each ray is refracted in an equal opposite 
direction to one another and so the net force imparted is 0. When displaced 
in the x-y plane the net force is unbalanced and points towards the centre 
of the trap; this principle can be generalised to 3 degrees of freedom if we 
consider a focused beam instead of a para-axial beam, typically the trap will be
weaker along the beam axis compared to the transverse trap strength.

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.65\linewidth]{ot_working_principal.png}
  \caption{Working principle for an optical trap, upper image shows a sphere at the centre of a trap experiences equal forces directed towards inwards. Lower image shows that when displaced 'ray a' is refracted far more than 'ray b' resulting in a net force back towards the centre.}
  \label{fig:working_principle}
\end{figure}

@@

These 4 equations describe how the electric and magnetic fields behave
and are related to one another at a microscopic level. Any discussion 
of optical trapping is underpinned by the fact that in all cases the 
Maxwell Equations must be satisfied at the macroscopic level, where 
one must account for the medium's interactions with the EM field. 
Macroscopic forms of the maxwell equations are given below \cite{Jackson_1975}:
\begin{align}
	\nabla \cdot \mathbf{D}
	&= \rho_f
	\\
	\nabla \cdot \mathbf{B}
	&= 0
	\\
	\nabla \times \mathbf{E}
	&= \frac{\partial \mathbf{B}}{\partial t}
	\\
	\nabla \times \mathbf{H}
	&= \mathbf{J}_f +\frac{\partial \mathbf{D}}{\partial t}  
\end{align}
Where $\rho_f$ and $\mathbf{J}_f$ are the charge density and current 
caused by the presence of free charges in the medium, and $\mathbf{D}$ 
and $\mathbf{H}$ are the displacement and magnetizing fields respectively. 
The latter two are defined by:
\begin{align}
	\nonumber
	\mathbf{D} &= \epsilon_0\mathbf{E}+\mathbf{P} 
	\\ 
	\nonumber
	\mathbf{H} &= \frac{1}{\mu_0}\mathbf{B}-\mathbf{M}
\end{align}

Where $\mathbf{P}$ is the polarization field and $\mathbf{M}$ is the 
magnetisation field, these fields arise due to the bound charges 
throughout the medium interacting with the EM field.

The force exerted by an optical tweezer can be subdivided into the 
gradient and scattering components, for most modelling research this 
is how the force fields are reported. The gradient force is a conservative 
force brought about by the polarisation of dielectric materials which 
is directed towards the point of maximum intensity (for a simple Gaussian 
beam this would be at it's focal point) \cite{YasuhiroHarada1996}. 
Typically, optical tweezers will utilise higher numerical aperture lenses 
in order to increase the intensity, thus allowing for stronger 
gradient forces.

The scattering force arises from the scattered field of the trapping beam 
pushing the target object away from the focal point. The equilibrium position 
(where the mean squared displacement is minimised) is found when the 
gradient force far exceeds the both the scattering force, and the thermal 
energy of the particle. Describing and calculating the optical forces is 
dependent on the ratio between the particle size and the trapping wavelength.

@@

The Ray-Optics model is the simplest to understand, the theory models 
the laser as a collection of individual 'rays' that propagate and are 
refracted according to Snell's Law. Based on the change in direction 
momentum is transferred to the target particle; with rays closest to 
the centre of the beam having greater intensity than those rays at 
the very edge of the beam. Consider a particle struck by two rays in 
a Gaussian beam (see Fig below), one coming close to the centre, and 
the other ray coming from the edge. As each ray is refracted by the 
target sphere a force is imparted onto it, the total force imparted 
is given by:
\begin{align}
	F_i = Q_i\frac{\Delta n P_i}{c}
\end{align}

where $Q_i$ is the trapping efficiency, $\Delta n$ is the difference
in refractive indices between the solution and the target particle,
and $P_i$ is the power of the individual ray. The net force can be 
subdivided into its gradient and scattering components, where the latter
directs the particle to the centre of the beam, and the latter acts on 
the particle in the direction of beam propagation. For a beam with a
Gaussian intensity distribution $P_i$ will fall off as you move from
the centre of the beam.

@@

Where $k$ is the wave number of the electromagnetic radiation ($k = 2\pi/\lambda$)
This combined with the constraints of Maxwell's Equations leaves 
very few exact solutions apart from spherical or planar waves, 
both of which can be converted between with relative ease. For 
example, a plane wave electric field can be expanded into spherical 
harmonics and likewise any spherical wave can be described as 
combination of plane waves offset from one another. For a single 
plane wave the incident, internal, and scattered fields are given as: 
\begin{align}
	\label{eq:incident}
	{\bf E}_{\rm inc}({\bf r})
	&=
    E_0 \sum_{nm}^\infty \left[
    a_{nm}{\bf M}_{nm}^{(1)}({\bf r})
    + b_{nm}{\bf N}_{nm}^{(1)}({\bf r})\right]
  \\
  \label{eq:internal}
  {\bf E}_{\rm int}({\bf r})
  &=
    E_0 \sum_{n=1}^\infty i^n\frac{2n+1}{n(n+1)}\left[
    -id_{nm}{\bf N}_{nm}^{(1)}({\bf r}) + c_{nm}{\bf M}_{nm}^{(1)}({\bf r})
    \right]
  \\
  \label{eq:scattered}
  {\bf E}_{\rm scat}({\bf r})
  &=
    E_0 \sum_{n=1}^\infty  i^n\frac{2n+1}{n(n+1)}\left[
      -i p_{nm}{\bf N}_{nm}^{(3)}({\bf r})+q_{nm}{\bf M}_{nm}^{(3)}({\bf r})
    \right] 
    \\
	\mathbf{E}_{tot}({\bf r}) &= 
	\begin{cases}
		{\bf E}_{\rm inc}({\bf r}) + {\bf E}_{\rm scat}({\bf r}) & \text{if \textbf{r} outside}
		\\ 
		{\bf E}_{\rm int}({\bf r}) & \text{if \textbf{r} inside}
	\end{cases}
\end{align}

where $a_{nm}$, $b_{nm}$, $c_{nm}$, $d_{nm}$, $p_{nm}$, and $q_{nm}$
are the expansion coefficients of each of the fields, and $M^{(1)}_{nm}$ 
and $N^{(1)}_{nm}$ are the vector spherical harmonics. For the incident
field computing its expansion coefficients is possible via analytical
methods and are completely dependent on the beam conditions imposed by
the user. However, computing the expansion coefficients of the
internal and scattered fields is far more tedious depending on the
shape of the target and what properties of the scattered field are
desired --- see sec.~\ref{sec:scattering}.  From an optical
trapping perspective the force exerted by a focused electric field can
be found by computing the Maxwell stress tensor which only requires
the total magnitude of the incident and scattered fields, essentially
computing the momentum contained in the incident beam and how much of
it has been transferred to the target particle.

Lorenz-Mie theory can be applied to describe the scattering from any 
particle regardless of size, though as the size of the target particle 
changes - relative to the light being scattered - the theory describing
the scattered light is better approximated by alternative theories. For 
particles that are far larger than the wavelength of light the scattering
is best described by the Ray-Optics theory. In the opposite case the 
target particle is better approximated as a single electric dipole as 
described by the Rayleigh scattering field.

@@

The Rayleigh approximation is the limiting approximation for describing 
a particles motion in an electromagnetic field who's wavelength is several 
times greater than the particle's size. The underlying theory is that a 
dielectric sphere can be treated as a dipole while in the presence of the 
electromagnetic field; in which case the scattering force is given simply 
by the scattering of the induced dipole, and the gradient force is due 
to the Lorentz force \cite{Gordon1973}. The gradient forces in the 
principle Cartesian axis are described by Harada et al \cite{YasuhiroHarada1996} 
in MKS units as a restorative rectangular force field:
\begin{align}
  F_{grad,x}
  &=-\hat{x} \frac{2\pi n_2 a^3}{c}
    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{x}/w_0}{1+(2\tilde{z})^2} \times I(r)
  \\
  F_{grad,y}
  &=-\hat{y} \frac{2\pi n_2 a^3}{c}
    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{y}/w_0}{1+(2\tilde{z})^2} \times I(r)
  \\
  F_{grad,z}
  &=-\hat{z} \frac{2\pi n_2 a^3}{c}
    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{y}/w_0}{1+(2\tilde{z})^2}
    \nonumber \times I(r)
  \\ 
  & \times \left[1-\frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right]
  \\
  \text{where:}
  \nonumber
  \\
	I(r) &= \left(\frac{2P}{\pi w_0^2}\right) \frac{1}{1+(2\tilde{z}^2)} 
	\exp \left[ - \frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right]
\end{align}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Scattering methods}
\label{sec:scattering}
There are several methods available to calculate the scattered field 
produced by a particle, while one can compute this from directly applying
Lorenz-Mie theory for complicated particle shapes and focused beams makes it 
tedious for some applications.  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{T-matrix Method}
The $T$-matrix method was first developed by Peter Waterman with his
research into acoustic wave scattering \cite{Waterman1969}, this 
would later be extended to electromagnetic waves. Sometimes being 
referred to as the extended boundary condition method (ECBM), 
the method replaces the scatterer with a series of surface currents
over the targets surface. These currents are chosen so that the
electric field outside is identical to the original problem \cite{Wriedt1998}. 
This choice of currents negates the need to compute the internal fields
which reduces the scattering problem to a system of linear equations 
relating the incident beam coefficients to the scattered beam coefficients
the relationship between each can be summarised as:
\begin{align}
	\label{eq:t_matrix}
	\begin{pmatrix}
		q_{mn} \\
		p_{mn} 
	\end{pmatrix}
	= \bold{T} 
	\begin{pmatrix}
		a_{mn} \\
		b_{mn}
	\end{pmatrix}
	= \begin{pmatrix}
		T_{11} \ T_{12} \\
		T_{21} \ T_{22}
	\end{pmatrix}
	\begin{pmatrix}
		a_{mn} \\
		b_{mn}
	\end{pmatrix}
\end{align}

Where $q_{mn}$, and $p_{mn}$ are the scattering beam coefficients, and
$a_{mn}$, and $b_{mn}$ are the coefficients of the incident beam. For 
this PhD in particular we utilise an extension of the ECBM known as 
multi-sphere $T$-matrix method (\textit{mstm}). Developed by Mackowski
\cite{Mackowski2011} the computational code computes the scattered field 
from each sphere within the target cluster and the incident field itself. 
The scattering between spheres slowly converges to a final result but modern 
codes truncate the calculations to an upper number of iterations (denoted by
$N_{max}$) to fit a desired accuracy and computational time.

The $T$-matrix method is exceptionally useful for computing the scattering 
from any arbitrary spherical aggregate. However, the $T$-matrix method by 
itself can be computationally taxing as the number of spheres increases 
\cite{Mackowski2011}. While it is possible to solve for the full electromagnetic 
field of each the entire cluster this is only applicable for a single 
orientation and can be even slower for large aggregates \cite{Mackowski1996, 
Xu1995}. The benefit of \textit{mstm} is that the major scattering properties 
(scattering and extinction cross sections, scattering matrices, etc) can 
all be computed both for single orientations, or averaged over multiple 
orientations to determine the average scattering from the target particle.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Discrete Dipole Approximation}
The discrete dipole approximation (DDA) is a general method that can 
be applied to the scattering from particles of arbitrary composition
and geometry. Developed by Purcell and Pennypacker \cite{Purcell1973}, 
the DDA method approximates the particle as being constructed of dipoles.
Each dipole interacts with both the incident field and the scattered fields
from every other dipole surrounding it. The resulting scattered field is 
identical to the scattered field produced by direct integration of 
Eq.~\eqref{eq:internal} throughout the full particle volume \cite{Goedecke1988}.
The integral form for the internal electric field inside a scatterer is 
given as:
\begin{align}
	\mathbf{E(r)} = \mathbf{E_{inc}(r)} + \int_{V/V_0}d^3r\mathbf{\bar{G}(r,r')}
	\chi(\mathbf{r'})\mathbf{E(r')}
\end{align}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Langevin Equation}
Describing any microscopic motion requires an understanding of a molecules
diffusive behaviour, for the case of optical tweezers the most complete model 
of diffusion is the Langevin equation. Models such as the Fickian, and Einstein 
derivations are sufficient for macroscopic behaviours the Langevin equation 
better describes the microscopic characteristics of any diffusive behaviour.

The Fickian model describes the net flux of solute molecules into a finite
volume of fluid being proportional to the density gradient $\delta\rho(u,t)/\delta u$
\cite{Gillespie2012}. It assumes that the solute molecules do not collide with 
one another or with molecules in the solution; overall, the Fickian model is
used to describe how solute molecules disperse over long periods of time. It 
does not provide any incite into the forces acting on individual molecules, 
whereas the Langevin model captures the physical interactions between an 
individual molecule and the surrounding fluid over a wide range of time scales.

The Einstein model of diffusion expands upon the Fickian model by considering
the collisions between individual molecules. If we consider a single particle 
suspended in a solution it will experience multiple collisions with other 
molecules \cite{Gillespie2012a} over a given time interval $\delta t$. The 
Einstein molecule allows us to consider how individual molecules behave while
suspended in a surrounding fluid. Where it begins to falter is when we consider
bringing $\delta t$ to $0$; in this scenario the molecule experiences not many
but only individual collisions. The Einstein model does not consider the inertia 
of the molecule in question and so for very short time frames the model implies 
that the molecule's velocity changes instantly after each collision \cite{Gillespie2012a,
Gillespie2012b}. Furthermore, the kinetic energy of each collision is not limited
by the thermal energy of the system, meaning that using the Einstein model to 
predict a particle's trajectory to an finite degree of accuracy implies that 
the particle has infinite kinetic energy \cite{Gillespie2012b}. This failure to 
describe motion over smaller time frames is addressed in the Langevin model by 
accounting for the fluid drag of the system, where any sudden change in velocity 
must result in a proportionally opposed drag force \cite{Gillespie2012c}.

The Langevin model of diffusion assumes that the net force on a particular 
particle is described fully by these individual collisions \cite{Gillespie2012c}. 
Unlike the Fickian model it provides a full description of the interactions 
between the target molecule and the surrounding fluid; while also being able
to describe its motion over any time scale, unlike the Einstein model:
\begin{align}
	m\frac{dv}{dt} + \gamma_0 v + F(t) = W(t)
\end{align}

@@

\eqref{eq:langevin} provides an accurate description of strongly trapped 
particles, however the analytical solution of the Langevin equation requires 
integration of the white noise term making it difficult to simulate the 
trajectory of a given particle \cite{Volpe2013}. Instead it is often far 
easier to solve the equation numerically and apply use the analytical 
solution to calibrate and extract information about the particle and fluid, 
and how the two interact with the optical trap.

\subsection{Finite Difference}
The Finite Difference approach involves discretizing the time and spatial 
elements in order to approximate the higher order terms. If we assume that 
$x(t)$ is differentiable to n (we can find its $n^{th}$ derivative) then 
we can use the Taylor series expansion to get:
\begin{align}
	x(t+\Delta t) = x(t)+\frac{x'(t)}{1!}\Delta t + \frac{x^2(t)}{2!}\Delta t^2+...+\frac{x^n(t)}{n!}\Delta t^n+R_n(x(t))	
\end{align}

Where $R_n(x(t))$ is the remainder term between the Taylor expansion to 
term n and the actual expression. If we limit our approach to the first 
derivative only we find that for sufficiently small values of $R_1$ the 
velocity and acceleration can be approximated as:
\begin{align}
	x'(t) &\approx \frac{x(t+\Delta t)-x(t)}{\Delta t}
	\\
	x^{''}(t) &\approx \frac{x'(t+\Delta t)-x'(t)}{\Delta t} = \frac{x(t)-2x(t+\Delta t)+x(t+2\Delta t)}{\Delta t^2}
\end{align}

By reversing the time step (i.e. use $-\Delta t$) to approximate the \\
velocity and acceleration based on the previous time steps we can 
discretize the position by taking finitely small  time steps (i.e. 
$x(t) = x_i,\ x(t-\Delta t) = x_{i-1}$). The same cannot be done for 
white noise as no information is known about $W(t)$ at any time. We 
can instead say that the velocity of a Brownian particle should 
approximate our white noise as a random walk, where at each new time 
step the position changes randomly within a given range. Constricting 
the variance to $\sqrt{2D}/\Delta t$ allows us to represent the white 
noise using the finite-difference approach as:
\begin{align}
	m\frac{x_i-2x_{i-1}+x_{i-2}}{\Delta t^2} = -\gamma\frac{x_i-x_{i-1}}{\Delta t}+\sqrt{2k_BT\gamma}\frac{w_i}{\sqrt{\Delta t}}
\end{align}

Where $w_i$ is a random real number between -1 and 1, we can say that 
it is normally distributed around 0 for simulation purposes. We can 
rearrange this for $x_i$ to approximate the Brownian motion of a 
particle (setting $x_0=0$), where the characteristic time is $\tau = m/\gamma$. 
Now in the case of an optical trap the restoration time scale is 
given by $\tau_{OT}=\kappa_x/\gamma$ which for strongly trapped particles 
is far greater than the characteristic time, therefore for simulation 
purposes we can neglect the particle's inertia which allows us to 
write the motion of an optically trapped particle as:
\begin{align}
	\label{eq:sim_langevin}
	x_i = x_{i-1} - \tau_{OT}\ x_{i-1}\Delta t + \sqrt{2D\Delta t}\ w_i
\end{align}
